Multivariate Density Estimation: Comparing Transformation Random Forest, Normalizing Flows and Copulas

### 2025-09-30 report/MSc_Report.Rnw
.

This thesis evaluates three approaches to multivariate density estimation for tabular data within a single, consistent pipeline: separable triangular transport maps (TTM-Sep), Transformation Random Forests (TRTF; additive predictor), and copulas (used only for low-dimensional diagnostics, $K\!\le\!3$). All methods use standardized inputs and a common evaluation protocol so that likelihoods, diagnostics, and compute are directly comparable. In the configuration studied (additive predictor and monotone CDF smoothing), TRTF and TTM-Sep yield the same triangular-likelihood form, which enables like-for-like evaluation.

>>>>

This thesis evaluates three approaches to multivariate density estimation for tabular data within a single, consistent pipeline: separable triangular transport maps (TTM-Sep), Transformation Random Forests (TRTF), and copulas. All methods use standardized inputs and a common evaluation protocol so that likelihoods, diagnostics, and compute are directly comparable. In the configuration studied, TRTF and TTM-Sep yield the same triangular-likelihood form, which enables like-for-like evaluation.
